Predicting Solid State Qubit Candidates

### 2021-11-25 paper.tex
Updates from Overleaf

\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
\usepackage{physics}
\usepackage{lipsum}
\usepackage{subfig}
% \usepackage{braket}

@@

\author{Oliver Lerstøl Hebnes}
\affiliation{Department of Physics and Center for Computing in Science Education, University of Oslo, N-0316 Oslo, Norway}
\author{Marianne Etzelm\"uller Bathen}
\affiliation{Advanced Power Semiconductor Laboratory, ETH Zürich, 8092  Zürich,  Switzerland}
\affiliation{Department of Physics and Center for Materials Science and Nanotechnology, University of Oslo, N-0316 Oslo, Norway}
\author{Øyvind Sigmundson Schøyen}
\affiliation{Department of Physics and Center for Computing in Science Education, University of Oslo, N-0316 Oslo, Norway}
\author{Sebastian G. Winther-Larsen}
\affiliation{Department of Physics and Center for Computing in Science Education, University of Oslo, N-0316 Oslo, Norway}
\author{Morten Hjorth-Jensen}
\affiliation{Facility for Rare Ion Beams and Department of Physics and Astronomy, Michigan State University, East Lansing, MI 48824, USA}
\affiliation{Department of Physics and Center for Computing in Science Education, University of Oslo, N-0316 Oslo, Norway}
\author{Lasse Vines}
\affiliation{Department of Physics and Center for Materials Science and Nanotechnology, University of Oslo, N-0316 Oslo, Norway}

@@

% Main message 2: Use method to propose new materials that are promising quantum hosts 
We find a lack of consistent results for the Ferrenti approach and the
augmented Ferrenti approach due to an overly broad formulation of the
training set, whereas the restrictions set in the insightful approach
proved suitable. All models agreed on $214$ predicted candidates, with
examples such as ZnGeP$_2$, MgSe, CdS, BP, BC$_2$N, BP, Ge, GeC, InP,
and InAs. All approaches and all models agreed on a subset of $47$
eligible candidates of $8$ elemental, $29$ binary, and $10$ tertiary
compounds.

@@

\section*{Introduction}%Oliver and Morten and Marianne (semiconductor portion)
To sustain the digital world’s increasing computational demand, alternatives to the classical computer must be explored. The quantum computer, together with the complementary quantum technologies of sensing and communication, is becoming a frontrunner in the race for solutions. 
Discuss recent benchmarks for quantum computing \cite{Arute_2019}.

Limitations of quantum computing and technology building blocks. Drawbacks of superconductors. Semiconductors are a promising platform. Example of NV center in diamond \cite{Doherty_2013}.

%From thesis - rewrite 
Unfortunately, there are substantial challenges associated with the modern quantum platforms simultaneously as the selection of quantum platforms are slim. The majority of discoveries of potential quantum platforms have so far happened by serendipity, and there is an urgent need for new and better materials that can escalate the effort for a sustainable future.

The fourth science paradigm introduces the potential of targeted search for promising material systems to act as qubit hosts. 
Herein, we provide a framework for the automatic discovery of promising solid-state material hosts using machine learning methods. 
Initial database building similar to Ferrenti et al \cite{Ferrenti2020}.

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{figures/ht-workflow.jpg}
    \caption{Placeholder for workflow. Morten: can you please add the other schematic figure we discussed as placeholder? Sebastian makes our own versions of these schematic figures (I can make some VESTA files with materials). }
    \label{fig:ht-workflow}
\end{figure}

\section*{Theoretical Background and Experimental Data} 
%Not all papers have a theory chapter, but since we are speaking to two quite different communities, it might be nice to include some basics. 
Quantum technologies, including sensing, communication and computing, 
are intriguing because of the new capabilities they offer. 
However, realization of large-scale fabrication for these techniques requires further technological development. 
The superconducting electronics platform that is currently employed for quantum computers necessitates millikelvin temperatures and may prove challenging to scale. 
Instead, recent developments have seen/heralded the rise of semiconductors as host materials for a variety of quantum devices. Much easier to combine with existing devices, and RT operation has been shown. 
However, the ideal host has yet to be identified, and the requirements for acting as a good host remain unclear. 
Different methods are needed.

\subsection*{Semiconductors as a Quantum Host} %Marianne and Lasse 
The key feature when considering a quantum object for application is coherence, which is again tightly connected to the concept of coherence. 
For all applications, the basic requirement is a two-level quantum system which is quite isolated so the levels remain stable. 
Also need to control the qubit state via initialization, manipulation and read-out. 
Requirements \cite{DiVincenzo2000,Weber2010}: 
\begin{itemize}
    \item A scalable physical system with well-characterized qubits
    \item Reliable qubit initialization protocols 
    \item Need quantum gates to operate on qubits 
    \item Qubit state coherence time must exceed gate operation time 
    \item Need to read out qubit state after operation. 
\end{itemize}

%Discuss experimental data and the semiconductor platform. 
Semiconductors embody point defects that may trap charge carriers, causing deep energy levels to appear within the fundamental band gap. 
Importantly, such deep levels can trap carriers in localized states that are essentially isolated from the surroundings, making them suitable for QT. 
A prime example is the nitrogen-vacancy center in diamond, which has been proposed 
as a contender for both qubit, sensing and communication applications \cite{Doherty_2013}. 
The NV center is a room temperature single-photon source emitting in the visible portion of the spectrum, exhibiting entangled photons that have been separated by ... km []. 
Moreover, the NV center in its negative charge state traps an electron in a localized and high-spin $S=1$ state that can be coherently controlled and with RT spin coherence of 1 ms []. 
Qubit operations and entanglement have been demonstrated []. 
However, emission wavelength is not ideally compatible with fiber optic technology (need telecom range), spin coherence is good but could be better, and sensing is good.

\begin{table}[t]
    \centering
    \begin{tabular}{c|c|c|c}
    Material & Band gap (eV) & Defect candidates & Reference \\
    \hline
    Diamond  & $5.5$  & N$_\mathrm{C}V_\mathrm{C}$, Si$_\mathrm{C}V_\mathrm{C}$, Ge$_\mathrm{C}V_\mathrm{C}$ & \cite{Balasubramanian_2009,Rogers_2014,Bhaskar_2018} \\  
    4H-SiC & $3.3$ & $V_\mathrm{Si}$, $V_\mathrm{Si}V_\mathrm{C}$, C$_\mathrm{Si}V_\mathrm{C}$, N$_\mathrm{C}V_\mathrm{Si}$ & \cite{Widmann2014,Christle_2015,Castelletto_2014,Zargaleh_2018} \\ 
    Si & $1.1$ & P, G, unidentified & \cite{Muhonen_2014,Durand_2020,Redjem2020} \\ 
    (2D) \textit{h}-BN & $6.0$ & Unidentified defects & \cite{Tran_2016,Tran_2016b,Hayee_2020} \\ 
    ZnO & $3.4$ & Unidentified defects & \cite{Morfa2012} \\ 
    ZnS & $3.6$ (zincblende) & Unidentified defects & \cite{Stewart2019} \\ 
    GaAs & $1.4$ & Quantum dots & \cite{Bluhm2010} \\ 
    GaN & $3.4$ & Quantum dots, unidentified defects & \cite{Roux2017,Berhane2018} \\
    AlN & $6.0$ & Unidentified defects & \cite{Xue2020}\\
    \end{tabular}
    \caption{Overview of materials and defects that have been demonstrated to exhibit QT-compatible characteristics such as single-photon emission and coherent spin manipulation.}
    \label{tab:qt-materials}
\end{table}

 ZnS [74], GaAs [75],
GaN [76, 77 , 80] and AlN [78, 79],  [81].

%QT requirements. 
%What parameters are we looking for? 
Table \ref{tab:qt-materials} contains an overview of known semiconductor materials with demonstrated quantum compatible characteristics. 
Not all defects are identified, and challenges due 
to the specific materials complicate the implementation of defects for QT. 
There are some common denominators for the materials: 
\begin{itemize}
    \item Wide band gap to facilitate localized deep levels (physical)
    \item Low spin orbit coupling to extend spin coherence times (physical) 
    \item Low el-phonon coupling for sharp ZPLs (physical) 
    \item Availability as high-quality, bulk, or thin-film single crystals (practical)
    \item Constituent elements with naturally occurring isotopes of zero nuclear spin (practical). 
\end{itemize}
Note the division between physical and practical requirements in the list above. 
The physical ones refer to some fundamental aspect of the material that permits specific properties to manifest, while practical refers to some more external influence that provides more easy measurement and detection of said properties. 
It is the former that we are interested in.

Different defect and material types may suit different needs. 
Therefore, we need a bigger bank of candidates. 
Also, we do not fully understand exactly what constitutes a good quantum host material. 
So far, detection of QT compatibility has been slow, as a signal is detected and then identified in a specific material. 
We are looking for a top-down approach to instead predict promising materials, and guide experiments. 
For this, we first need a large dataset.

\subsection*{Density Functional Theory Output} %Marianne 
Density functional theory has been exceptionally successful in predicting solid-state material properties 
such as

Data formats, limitations etc.

Density functional theory is a versatile and powerful theoretical approach, enabling targeted study of semiconductor materials with many atoms. 
Proven successful for material properties such as thermal stability, lattice parameters and band gap.

\subsection*{Material Informatics} %Øyvind (Oliver)    
%Data mining, screening

\subsection*{Machine Learning} % Morten

\section*{Results}
Includes discussion of results.

\subsection*{Information Flow}  %Sebastian/Oyvind (Marianne) 
Includes results from data mining and description of the three approaches. 
%MEB: Since Methods are at the end, I propose to put part of this discussion in the Results section. 
%This is because it is an important result of the work, and also because it is important to understand the rest of the paper

\subsection*{Optimization of Machine Learning Models} %Morten (Oliver input in round 2)

\subsection*{Predicting Novel Material Hosts for Quantum Technology} %Oyvind (Oliver input in round 2) 
\subsubsection*{The Ferrenti approach} 
\subsubsection*{The augmented Ferrenti approach}
\subsubsection*{The insightful approach}
\subsubsection*{Comparing the approaches}

\section*{Discussion} % All - last thing we write  
Shorter section, includes summary, concluding remarks and is quite heavy on outlook.

\section*{Methods}
%Methodology comes after concluding remarks, and is structured into subsections

\subsection*{Databases} %Oliver and Øyvind 
%Materials project, AFLOW, Open Quantum, JARVIS 
Mining.

\subsection*{Material Informatics} %Oliver and Øyvind 
Screening and workflow and approaches.

\subsection*{Machine Learning} %Sebastian/Morten   
Insight gaining.

\section*{Data availability} 
%We can put something else, standard text 
The data that support the plots within this paper and other findings of this study are
available from the corresponding authors upon reasonable request.

\section*{Code availability} 
The codes developed in this study are available from the authors upon reasonable
request.

>>>>

\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
\usepackage{physics}
\usepackage{lipsum}
% \usepackage{subfig} %not compatible with subcaption
% \usepackage{braket}
\usepackage{siunitx}
\usepackage{color}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{tikz,pgfplots}
\usepackage{makecell}
\captionsetup[table]{skip=10pt}
\pgfplotsset{compat=1.17}
\sisetup{separate-uncertainty}
\usepackage{hyperref} 
\usepackage{longtable}

\usepackage{forest}
\forestset{
    .style={
        for tree={
            base=bottom,
            child anchor=north,
            align=center,
            s sep+=1cm,
    straight edge/.style={
        edge path={\noexpand\path[\forestoption{edge},thick,-{Latex}]
        (!u.parent anchor) -- (.child anchor);}
    },
    if n children={0}
        {tier=word, draw, thick, rectangle}
        {draw, diamond, thick, aspect=2},
    if n=1{%
        edge path={\noexpand\path[\forestoption{edge},thick,-{Latex}]
        (!u.parent anchor) -| (.child anchor) node[pos=.2, above] {Y};}
        }{
        edge path={\noexpand\path[\forestoption{edge},thick,-{Latex}]
        (!u.parent anchor) -| (.child anchor) node[pos=.2, above] {N};}
        }
        }
    }
}
\tikzset{
  green arrow/.style={
    midway,green,sloped,fill, minimum height=2cm, single arrow, single arrow head extend=.5cm, single arrow head indent=.25cm,xscale=0.3,yscale=0.15,
    allow upside down
  },
  yellow arrow/.style={
    midway,yellow,sloped,fill, minimum height=2cm, single arrow, single arrow head extend=.5cm, single arrow head indent=.25cm,xscale=0.3,yscale=0.15,
    allow upside down
  },
  black arrow/.style 2 args={-stealth, shorten >=#1, shorten <=#2},
  black arrow/.default={1mm}{1mm},
  tree box/.style={draw, rounded corners, inner sep=1em},
  node box/.style={white, draw=white, text=black, rectangle, rounded corners},
}

@@

\author{Oliver Lerstøl Hebnes}
\affiliation{Department of Physics and Center for Computing in Science Education, University of Oslo, N-0316 Oslo, Norway}
\affiliation{Sopra Steria, N-0185 Oslo, Norway}

\author{Marianne Etzelm\"uller Bathen}
\affiliation{Advanced Power Semiconductor Laboratory, ETH Zürich, 8092  Zürich,  Switzerland}
\affiliation{Department of Physics and Center for Materials Science and Nanotechnology, University of Oslo, N-0316 Oslo, Norway}

\author{Sebastian G. Winther-Larsen}
\affiliation{Menon Economics, N-0369 Oslo, Norway}
\affiliation{Department of Physics and Center for Computing in Science Education, University of Oslo, N-0316 Oslo, Norway}

\author{Øyvind Sigmundson Schøyen}
\affiliation{Department of Physics and Center for Computing in Science Education, University of Oslo, N-0316 Oslo, Norway}

\author{Morten Hjorth-Jensen}
\affiliation{Department of Physics and Astronomy and Facility for Rare Ion Beams, Michigan State University, East Lansing, MI 48824, USA}
\affiliation{Department of Physics and Center for Computing in Science Education, University of Oslo, N-0316 Oslo, Norway}

\author{Lasse Vines}
\affiliation{Department of Physics and Center for Materials Science and Nanotechnology, University of Oslo, N-0316 Oslo, Norway}

@@

% Main message 2: Use method to propose new materials that are promising quantum hosts 
We find a lack of consistent results for the Ferrenti approach and the augmented
Ferrenti approach due to an overly broad formulation of the training
set, whereas the restrictions set in the insightful approach proved suitable. All
models for the insightful approach agreed on $214$ predicted candidates, with examples such as ZnGeP$_2$, 
MgSe, CdS, BP, BC2N, BP, Ge, GeC, InP, and InAs. All three approaches and all ML
models agreed on a subset of $47$ eligible candidates of $8$ elemental, $29$ binary,
and $10$ tertiary compounds.

@@

\section*{Introduction}%Oliver and Morten and Marianne (semiconductor portion)
The second quantum revolution is heralding the development of new solutions and devices based on quantum technology \cite{Acin2018}. Among the promising prototypes that are already available we find in vivo sensing of magnetic fields in cells [], secure communication over large distances by separation of entangled photons [], and finally the demonstration of quantum computing []. 
Quantum computers are in high demand to meet the increasing need for computing power to solve complex many-body problems that are beyond the scope of classical machines.   
Simultaneously, the concepts of entanglement and teleportation may enable advanced quantum communication protocols such as the quantum internet. 
Recent advances have highlighted the potential of quantum computer architectures to outperform classical systems as a 53-qubit quantum computer based on superconducting electronics solved a computational problem that was beyond the capabilities of a 200'000 core supercomputer \cite{Arute_2019}, spurring further investigations into competitive quantum technology platforms.  And IBM announced recently its 127 qubit quantum processor.

%Figure 1 introduction schematic 
\begin{figure}[t]
    \centering
    %\includegraphics[width=0.6\textwidth]{figures/paradigms.png}
    \includegraphics[width=0.9\textwidth]{figures/ht-workflow.jpg}
    \caption{[PLACEHOLDER] Schematic of an example workflow in material informatics. 
    %\marianne{Morten: can you please add the other schematic figure we discussed as placeholder? Sebastian makes our own versions of these schematic figures (I can make some VESTA files with materials). We should have just one figure I think, perhaps even combining the paradigms and explaining the material informatics worklow.} 
    }
    \label{fig:paradigm}
\end{figure}

Several platforms are available for the development of quantum technologies (QT) but the materials and fabrication technologies are less mature than those for, e.g., classical computers and sensors. 
An important concern in this context is that of scalability. 
For example, the best performing quantum computer prototypes available today rely on superconducting electronics or trapped ions. Both technologies require millikelvin temperatures to operate, fabrication is challenging, and the stability of interactions between qubits is an issue. Instead, semiconductors are emerging as a promising alternative platform, offering competitive characteristics combined with the possibility of room temperature operation and mature and scalable material processing and fabrication.

Quantum technologies based on semiconductors rely on either defects or quantum dots where the latter kind can be of the self-assembled or fabricated type. 
Semiconductor defects can act as single-photon emitters or spin centers and are compatible with the three main QT categories of computing, communication and sensing.  
These characteristics are most often found for the case of defects that introduce deep energy levels into the semiconductor band gap. So-called deep level defects can trap carriers in localized states that are essentially isolated from the surroundings, making them highly suitable for QT due to, e.g., reliable single-photon emission and long spin coherence times. 
The most well-known QT defect is the negatively charged nitrogen-vacancy (NV) center in diamond \cite{Doherty_2013}, but silicon carbide (SiC) and the various quantum emitters therein are poised to out-compete diamond in certain areas due to the favorable emission wavelength region in the near infrared and telecom coupled with more mature material processing and fabrication  \cite{Bathen2021}. 
However, defect based QT is still in the early stages, and the issues left to address include identification of suitable host materials and candidate defects, and scalable and reproducible quantum device fabrication. Another important issue is that the understanding of the requirements for a material to enable quantum compatible properties to manifest remains lacking,  
%The plethora of possible emitters and spin centers formed by semiconductor defects or nanostructures come with different characteristics that are compatible with specific applications.  
%Additionally, the understanding of what exactly makes a semiconductor based quantum platform suitable for QT is scarce, 
and the selection of quantum compatible semiconductor host materials is slim.

The majority of discoveries of quantum emitters in semiconductors has so far happened by serendipity, and there is an urgent need for 
%new and better materials that can escalate the effort for a sustainable future. 
a better understanding of how different materials properties manifest. 
In this context, a framework for dedicated materials search and analysis is needed. 
The fourth science paradigm of big data driven science presents a significant improvement over the third paradigm focusing on computational science, for instance via density functional theory calculations of different materials properties. The fourth paradigm 
%, as illustrated by the schematic in Figure~\ref{fig:paradigm}(a), 
introduces the potential of targeted search for promising material systems to act as hosts for qubit and single-photon emitter defects.  
Instead of searching through a host of signals for the one that matches our criteria, we are now able to \textit{predict} which materials and signatures we should target for more detailed studies. 
This is made feasible by the availability of databases containing material properties for a wide range of different systems. In this work, the data in question is provided by bulk density functional theory (DFT) calculations to obtain the ground state properties of different elements and compounds. Combined with machine learning algorithms we provide a path towards precise classification of qubit candidates. The inclusions of machine learning methods follows recent trends in applications of 
statistics, data science and machine learning for scientific discoveries, see for example Refs.~\cite{deiana2021,Carleo2019}.

In this work we provide a framework for the data mining and automated discovery of promising solid-state material hosts for quantum emitters and spin centers using targeted database search and machine learning methods combined with knowledge from the field. The methodology developed herein can be modified for other material types and application areas provided that a large enough database with relevant theoretical or experimental data is available. 
The method that was developed relies on data extraction from the databases Materials Project \cite{Jain2013,Jain2018}, The Open Quantum Materials Database (OQMD) \cite{Saal2013, Kirklin2015}, JARVIS-DFT \cite{Choudhary2020}, AFLOW \cite{Curtarolo2012, Curtarolo2012a, Calderon2015} and AFLOW-ML \cite{Isayev2017}. 
Matminer \cite{Ward2018} is then used for material analysis to featurize the extracted data. 
An important feature of the work is the database building and development of the training sets for the machine learning models. We have developed three different approaches to data mining: (i) \emph{the Ferrenti approach} which is similar to that proposed by \citeauthor{Ferrenti2020}  \cite{Ferrenti2020}, (ii) \emph{the augmented Ferrenti approach} and (iii) \emph{the insightful approach}. We find that the former two methodologies base their data extraction protocols on practical considerations, such as experimental availability in specific forms or the toxicity of the materials, leading to few observable trends in the materials predicted by well-tested machine learning algorithms, such as logistic regression, decision trees, and ensemble methods like random forests and gradient boosting \cite{mehta2019,hastie2009}. The third approach, on the other hand, relies on materials with experimentally proven advantageous characteristics in the training set. This more restricted approach proved more suitable than the other two and yielded $214$ predicted candidates with examples such as ZnGeP$_2$, MgSe, CdS, BP, BC$_2$N, BP, Ge, GeC, InP, and InAs. 
%\begin{itemize}
 %   \item Data was extracted from databases materials project ... 
 %   \item Featurization using matminer 
 %   \item Initial database building similar to Ferrenti et al \cite{Ferrenti2020}, but method is expanded to focus mainly on physical aspects. 
 %   \item Machine learning on top.  
%\end{itemize}

%\section*{Theoretical Background and Experimental Data}

%Quantum technologies, including sensing, communication and computing, are intriguing because of the new capabilities they offer. However, realization of large-scale fabrication for these techniques requires further technological development. The superconducting electronics platform that is currently employed for quantum computers necessitates millikelvin temperatures and may prove challenging to scale. Instead, recent developments have seen/heralded the rise of semiconductors as possible host materials for a variety of quantum devices. Semiconductor quantum dots or point defects could prove easier to combine with existing devices, and room temperature operation of prototype communication and computing devices has been shown.  However, this technology is still in the early stages, and the issues left to address include scalable and reproducible device fabrication, identification of ideally suited host candidates, and, importantly, understanding the requirements for a material to act as a good host remain unclear.

%\subsection*{Semiconductors as a Quantum Host} 
%Quantum technologies commonly rely on a quantum bit or qubit as the basis for operation. \marianne{Some schematic is shown in Figure~\ref{fig:qt}}. For all applications, the basic requirement is a two-level quantum system which is essentially isolated from its environment, ensuring sufficient stability to perform various operation. In other words, we search for systems which exhibit a high degree of coherence. Additionally, the qubit state must be controlable via repeatable initialization, manipulation and read-out protocols. A set of requirements were formulated by DiVincenzo for the case of qubits and quantum computing \cite{DiVincenzo2000}. Below, these are rephrased to encompass qubit requirements for all three QT application areas:  
%\begin{itemize}
 %   \item A scalable physical system with well-characterized qubits
 %  \item Reliable qubit initialization protocols 
 %   \item Quantum gates and protocols to operate on and control qubit states  
 %   \item Qubit state coherence time must exceed gate operation time 
 %   \item Qubit state read-out after operation, preferably optically.  
%\end{itemize}

%\begin{figure}[t]
 %   \centering
 %   \includegraphics[width=0.3\textwidth]{figures/qubit.png}
 %   \caption{[PLACEHOLDER] Schematic of classical versus quantum bits. }
 %   \label{fig:qt}
%\end{figure}

%Discuss experimental data and the semiconductor platform. 
%Semiconductors embody point defects that may trap charge carriers, causing deep energy levels to appear within the fundamental band gap. Importantly, deep level defects can trap carriers in localized states that are essentially isolated from the surroundings, making them suitable for QT. A prime example is the nitrogen-vacancy (NV) center in diamond, which has been proposed as a contender for both computing, sensing and communication applications \cite{Doherty_2013}. The NV center is a room temperature single-photon source emitting in the visible portion of the spectrum, and has been employed for, e.g., stress tensor mapping \cite{Broadway2019} and optical imaging of living cells \cite{Lesage_2013}. Moreover, the NV center in its negative charge state traps an electron in a localized and high-spin $S=1$ state that can be coherently controlled and with RT spin coherence times above \SI{1}{\milli\second} \cite{Doherty_2013}, leading to demonstrations of entanglement and teleportation of two NV center spins \cite{Bernien2013,Pfaff_2014}. However, despite these promising discoveries, diamond  as a quantum platform comes with several important drawbacks. The relevant defects therein, such as the NV center, emit in the visible region of the spectrum which is not compatible with the telecom wavelengths needed to integrate with optical fiber technology. Additionally, diamond is expensive, device fabrication is challenging and scalability is poor, leading to concerns in whether quantum technology device development will be realizable on this platform. Hence, we search among the plethora of available semiconductors for a platform that can deliver on the necessary parameters for QT applications while also, ideally, providing availability and scalable fabrication. In this context, the material properties that facilitate quantum compatible characteristics such as single-photon emission and coherent spin manipulation to manifest must be identified.

%\begin{figure}[t]
%    \centering
%    \includegraphics[width=0.5\textwidth]{figures/NV.png}
%    \caption{[PLACEHOLDER] Schematic representation of the electronic structure of the NV center in diamond.}
%    \label{fig:defects-dft}
%\end{figure}

%\begin{figure}[t]
 %   \centering
 %   \includegraphics[width=0.5\textwidth]{figures/defects-sic.pdf}
 %   \caption{[PLACEHOLDER] Schematic representation of different quantum compatible defects in SiC and the energy levels they introduce into the band gap, from Ref.\cite{Bathen2021}.} 
 %   \label{fig:defects-sic}
%\end{figure}

%Table \ref{tab:qt-materials} contains an overview of known semiconductor materials with demonstrated quantum compatible characteristics. In this context we consider single-photon emission and coherent spin manipulation for systems related to point defects primarily, but also self-assembled or lithographically structured quantum dots. \marianne{A (placeholder) schematic of quantum compatible defects and the energy levels they introduce into the band gap of 4H-SiC is shown in Figure~\ref{fig:defects-sic}.} Most of the defects in the table have yet to be identified, and different material related challenges complicate the implementation of defects for QT. A set of common denominators for materials that contain qubit defects was summarized by Weber et al. \cite{Weber2010}: 
%\begin{itemize}
 %   \item Wide band gap to facilitate localized deep levels (physical)
 %    \item Low spin orbit coupling to extend spin coherence times (physical) 
 %   \item Low el-phonon coupling for sharp emission lines (physical) 
 %   \item Availability as high-quality, bulk, or thin-film single crystals (practical)
 %   \item Constituent elements with naturally occurring isotopes of zero nuclear spin (practical). 
%\end{itemize}
%Note the division between physical and practical requirements in the list above. The physical criteria refer to some fundamental aspect of the material that permits specific properties to manifest, while practical refers to the external influences that simplify measurement and detection of the interesting properties. In this work, we will focus primarily on the former set of criteria, as understanding them better may give further insights into the makings of a reliable QT platform.

%In addition to obtaining greater understanding of  exactly what constitutes a good quantum host material, we must recognize that different defect and material types may suit different needs. Therefore, we aim to develop a bigger bank of candidate defects and materials. So far, detection of QT compatibility in semiconductors has been rather slow, depending on signal detection and later identification in each specific material. Instead, we are looking for a top-down approach to \textit{predict} promising materials, enabling a guiding of experimental efforts.

%\subsection*{Density Functional Theory Output} %Marianne 
%The density functional theory (DFT) is a versatile and powerful theoretical approach that has been exceptionally successful in predicting solid-state material properties such as stability of different phases, lattice parameters and band gaps. DFT provides a method for iteratively solving the Schr{\"o}dinger equation based on initial guesses for the atomic potentials. Using DFT calculations, we are able to accurately investigate semiconductor materials with many atoms while keeping efficiency and cost in mind.

%The fundamental challenge of the DFT is that the calculations rely on approximations to the potential function of a system as expressed by the exchange-correlation (XC) functional, which is not known for bulk materials. The XC functional is usually grouped into three categories based on the level of theory. The local density approximation (LDA) takes the homogeneous electron gas as a starting point, while the more advanced general gradient approximation (GGA) includes also the gradient in local electron gas variations. These local or semi-local functionals yield satisfactory results on geometrical concerns but tend to underestimate energies such as the band gap or defect energy levels due to a self-interaction error (i.e., an electron is allowed to interact with itself) [Ref SchollSteckel] \cite{Freysoldt2014}. More recent developments have seen the widespread use of hybrid functionals such as the HSE06 incorporating a portion of Hartree-Fock exact exchange \cite{Heyd2003},  albeit at a much greater computational cost.

%Despite the limitations imposed by the approximate nature of the XC functional, and the band gap error of LDA and GGA level functionals, DFT can be a powerful tool when predicting, designing and interpreting experiments. Herein, we employ databases containing DFT calculations of material properties to gain insight into the fundamental mechanisms behind observed quantum characteristics, and predict new candidate QT hosts. The calculations we employ were performed using the GGA-level Perdew-Burke-Erznerhof (PBE) functional \cite{Perdew1996}. Note that the PBE functional tends to underestimate the band gap of semiconductor materials. Nonetheless, a larger data set is available than for, e.g., HSE06, and many trends are conserved also with PBE. The ground-state (and $0$~K) parameters that can be extracted from this type of data set include the total energy (translates into stability), the band gap, the electronegativity, the electron-phonon coupling and magnetic properties.

%\begin{figure}[t]
%    \centering
%    \includegraphics[width=\textwidth]{figures/ht-workflow.jpg}
%    \caption{[PLACEHOLDER for workflow] Morten: can you please add the other schematic figure we discussed as placeholder? Sebastian makes our own versions of these schematic figures (I can make some VESTA files with materials). }
%    \label{fig:ht-workflow}
%\end{figure}

%\subsection*{Machine Learning}

%Should contain a brief description of the four ``algorithms'':
%\begin{itemize}
 %   \item Logistic regression
 %   \item Decision trees
 %   \item Random forest
 %   \item Gradient boost
%\end{itemize}

%Why these? Works well out of the box, I guess. Readily available in scikit-learn.

\section*{Results and discussion}
%Includes discussion of results.

\subsection*{Information Flow}  %Sebastian/Oliver (Marianne) 
% MEB: Since Methods are at the end, I propose to put part of this
% discussion in the Results section. 
% This is because it is an important result of the work, and also
% because it is important to understand the rest of the paper

The information stream in this work can be regarded as many
modular parts connected in logical pieces. The initial step for
gathering material data and building features is visualized by the
the flowchart in Figure~\ref{fig:flowchart-makedata}. 
Initially, we start by extracting all entries in
the Materials Project that matches a specific query.

\input{figures/tikz-plots/flow-chart-data-extraction.tex}

\input{figures/tikz-plots/flow-chart-featurization.tex}

Thereafter, we apply Matminer’s \cite{Ward2018}
featurization tools to make thousands of features of the data. The conditions for this initial query are that the materials must derive from experimental data, and have a band gap wider than $0.1$~eV. In a parallel step, entries that are deemed
similar to the entries from the initial query
are extracted from 
OQMD \cite{Saal2013,Kirklin2015}, JARVIS-DFT \cite{Choudhary2020}, 
AFLOW \cite{Curtarolo2012, Curtarolo2012a, Calderon2015}, 
AFLOW-ML \cite{Isayev2017},
and
Citrine Informatics REF!??. The results of these steps are combined into
a data set for further analysis.

A schematic describing the featurization process in Matminer is shown in Figure~\ref{fig:flowchart-featurization}. 
\textbf{[Describe it and say something about which features we choose.]}  
The 39 features selected in this work are summarized in the Supplementary Material at \cite{supplementary}.

\subsection*{Data Mining}

After compiling a dataset from the procedure outlined above, we 
face a challenge in terms of defining a training set that %we can train data on. 
can be used to train an ML algorithm. 
This is challenging for two reasons. 
Firstly, although defects like the NV center in diamond and emitters in SiC are becoming
increasingly well studied, few candidate materials are known. An additional consideration
is that the physical mechanisms promoting favorable defect properties are not fully understood. 
Secondly, defining materials as unsuitable candidates for QT has proven to be surprisingly
intricate. 
%This is not only challenging due to the lack of known suitable candidates (1),  but also due to the intricacy of defining materials as unsuitable candidates (0). 
We follow three separate procedures to define a training set consisting of both suitable and unsuitable material candidates for QT.

\subsubsection*{The Ferrenti approach}

The first approach to defining a training set is based on the criteria
from the paper by \citeauthor{Ferrenti2020} \cite{Ferrenti2020}.
They suggest a data mining process consisting of four stages by 
systematically evaluating the suitability of host materials from
the Materials Project.

In this framework, we label \textbf{suitable} candidates by the following steps. 
\begin{enumerate}
    \item Include materials that;
    \begin{itemize}
        \item contain elements with more than $50\%$ natural abundance of 
            zero spin isotopes,
        \item crystallize in non-polar space groups,
        \item are present in the ICSD database, and
        \item are calculated to be nonmagnetic. 
    \end{itemize}
    \item Pragmatically remove toxic, radioactive and otherwise 
        ``difficult'' materials;
    \begin{itemize}
        \item exclude Th, U, Cd and Hg,
        \item exclude any rare-earth metals and noble gases,
        \item exclude Ru, Os, Fe or Ni.
    \end{itemize}
    \item Include only materials with a band gap larger than $0.5$~eV calculated using DFT and the PBE functional. 
    \item Ensure that energy above hull is less than $0.2$~eV per atom.
\end{enumerate}

We then label \textbf{unsuitable} candidates as materials in the ICSD database that crystallize in polar space groups, are calculated 
to be magnetic and have a band gap larger than $0.1$~eV as calculated by PBE-GGA.

\subsubsection*{The augmented Ferrenti approach}
In the second approach, we adjust the first approach in order to 
improve the dataset. This approach is therefore named
\emph{the augmented Ferrenti approach}.

In the first approach we included criteria that were not physically motivated, but instead based on a purely practical perspective, like excluding toxic and
radioactive elements. Such criteria are not necessarily related to what makes 
a material suitable for quantum technology - eventually it is up to the 
experimentalist to decide the practicality of the material to be employed. 
Consequently, in the augmented Ferrenti approach, we remove stage two from the approach above. What is more, we will include a few additional elements that have shown promising properties, but were initially excluded due to the lack of spin zero isotopes.

The following steps make up the process of choosing \textbf{suitable} candidates 
in the \emph{augmented Ferrenti approach},
\begin{enumerate}
    \item Include materials that; 
    \begin{itemize}
        \item contain elements where more than half have a natural abundance of zero spin isotopes, except Al, P, Ga, As, B and N, 
        \item crystallize in non-polar space groups,
        \item are present in the ICSD database,
        \item are calculated to be nonmagnetic. 
    \end{itemize}
    \item Only keep materials that have a band gap larger than $1.5$~eV as calculated by PBE-GGA,
    \item Ensure energy above hull is less than $0.2$~eV per atom. 
\end{enumerate}

For \textbf{unsuitable} candidates, we implement the same strategy as the non-augmented
Ferrenti approach. The result is a somewhat unbalanced dataset, with up to $75 \ \%$ of the 
materials found in the suitable group. However, the dataset is $78 \ \%$ larger than for
the Ferrenti approach.

\subsubsection*{The insightful approach}
The third approach differs substantially from the first two approaches in terms 
of labeling, as we apply knowledge from the field (see, e.g., Refs.~\cite{Atatuere2018,Zhang2020,Son2020,Toth2019,Bathen2021} for an overview)   
to pick our promising material host candidates. We therefore name this scheme 
\emph{the insightful approach}. 
Due to the concern of having a too-small dataset, we will include
materials that are promising and have shown suitable properties to 
accommodate deep defects that potentially can exhibit quantum effects. 
Materials where properties like single-photon emission and spin manipulation are attributed to quantum dot self-assembled formation or lithographic structuring will also be included in this context.  
The approach to pick \textbf{suitable} candidates can be summed up as;   
\begin{enumerate}
    \item Pick candidates that match the formulas in Table~\ref{tab:qt-materials}, or the formulas ZnSe,  AlP, GaP, AlAs, ZnTe, CdS \cite{Weber2010} or SiGe \cite{Hardy2019}. 
    \item Make sure the candidates are present in the ICSD database.  
    \item Perform a manual screening for correct structures.
\end{enumerate}
